---
title: "Accessing spatial data for Missouri"
toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background
Most of the spatial data in the MDCSpatialData package are derived from two R packages, **tigris** and **FedData**. Tigris lets you download sf objects of files from the national TIGER data set, while FedData lets you download spatial data from a variety of sources (NHD, NLCD, NASS, SSUGO, etc.) Your best bet is to look at those reference manuals, but I give a few quick examples here.

### Plotting the Missouri spatial boundaries

This can be done easily with ggplot
```{r,eval=FALSE}

#not run

MDCSpatialData::counties.mo->counties.mo
MDCSpatialData::state.mo->state.mo

ggplot(counties.mo)+
  geom_sf()

ggplot(state.mo)+
  geom_sf()
```


### Extracting spatial data from **tigris**

```{r,eval=FALSE}
#not run
tigris::counties(state='MO')->counties.mo # census blocks

tigris::blocks(state=29))->blocks.mo


# school district boundaries
# see tigris documentation

tigris::school_districts(state=29)->schools.mo


```

### Extracting spatial data from **FedData**

There are a number of spatial, landcover, and weather data sets you can pull from. Definitley read the reference manual for this package to get a comprehensive idea of what data are available.

```{r,eval=FALSE}
# need a boundary shape. If it's not a rectangle or a square, one will be created internal to the function before extracting the data. The bounding area will be extracted from the min/max X and Y coordinates of the polygon.



#get NASS Cropland data layer data this has a size limit for the raster

MDCSpatialData::countes.mo|>
  dplyr::filter(NAME=='Boone')->boone.co
FedData::get_nass_cdl(template=boone.co,
                 label='MO',
                 year=2023)->cdl.boone

terra::plot(cdl.boone)
# returns a SpatRaster, a terra raster object class


#get NHD data
FedData::get_nhd(template=MDCSpatialData::state.mo,
                 label='MO',   #for use if writing the data to file vs storing in memory.
                 nhdplus=TRUE)->nhd.mo

#Returns a list with 5 sf objects: Points,lines, flowlines, areas (streams and lakes), and water bodies.

# get HUC12 data (the only scale available in FedData)

FedData::get_wbd(template=MDCSpatialData::state.mo,label='MO')->huc.mo

#returns a sf object

```



